---
title: "Re-visiting Obesity Gender Gaps: New insights for policy under a time-bound perspective"
author: "Aguirre Hörmann, Pablo, Saravia-Matus, Silvia, Morales Hermosilla, Valentina, and Berdegué, Julio"
output: 
  word_document:
    reference_docx: ObesityPaper_Format.docx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, comment = FALSE, cache = TRUE)
```

```{r}
source("0.ObesityGAP_Packages.R", encoding = "UTF-8")
source("1.ObesityGAP_LoadData.R", encoding = "UTF-8")
source("2.ObesityGAP_DataWrangling.R", encoding = "UTF-8")
```

#### Abstract:

#### Keywords:

#### JEL Classifications:

# 1. Introduction

Higher prevalence of obesity among women than men, particularly in low and middle-income economies with high gender inequality indices has been documented in several studies (Garawi, Devries, Thorogood, & Uauy, 2014; Wells, Marphatia, Cole, & McCoy, 2012 ; Brooks & Maklakov, 2010; Low & Binns, 2015; Williams, Mesidor, Winters, Dubbert, & Wyatt, 2015). For instance, the difference of the prevalence of obesity between female and male in 1990 was 3.7 percentage points, but by 2016, this difference had increased by 18% (Authors calculations based on data from the WHO (2019). When aiming to provide an explanation for this phenomenon, cross sectional analyses have indicated that gender inequality is one if not the main root cause (Garawi et al, 2014; Ferretti & Mariani, 2017; Wells, J. C. K., Marphatia, A. A., Cole, T. J., & McCoy, D. (2012) ). However, it is not clear what would be the key entry points for policy action given that gender inequality touches on a range of socio-economic variables and environments. Simultaneously, studies incorporating time trends to analyze obesity have been mainly descriptive in nature and not necessarily focused on identifying the triggers of increasing gender gaps in excess malnutrition (Bentham et al., 2017;Fallah-Fini, Rahmandad, Huang, Bures, & Glass, 2014). 
 
Therefore, the aim of this paper is to explore drivers behind the gender gap in the prevalence of obesity at a country-level using data from 1995 (first year available for the gender inequality index) to 2016 (last year available of obesity data). The idea is to better understand the evolution of gender differences in excess malnutrition and identify relevant policy entry points. The novelty in comparison to previous studies is to address this phenomenon under a panel data regression controlling for both country and time effects while also considering the different (policy –related) elements behind gender inequality. In addition, we  tested the different estimated models against  their predictive accuracy in order to dig further into the relevance of our findings and also compare these models with other estimations within the realm of machine learning techniques aiming to increase predictive power for future gender gaps in obesity. 

The paper is structured as follows: Section 2 begins with a brief description of literature findings and an overview of how gender gaps in obesity have evolved at world and regional levels. Section 3 presents the dataset and methodological approach. We explore under a panel dataset with global coverage, the association between gender inequality (using UNDP’s Gender Inequality Index) and other frequently-reported control variables such as, income, smoking, urbanization, fertility rates, etc. (Althoff et al., 2017; Brooks and Maklakov, 2010). The functional forms used to estimate the determinants of obesity gender gaps are drawn from Garawi et. al (2014) and these specifications are used as the basis for decomposing the gender inequality index into its three main dimensions: Health, Empowerment, and Labor Market Participation.  Section 4 tests the predictive capabilities of the proposed explanatory model as well as other specifications drawn from machine learning literature, while section 5 concludes.

# 2. Obesity Gender Gap Literature and Trends 1995-2016


# 3. Methods and Results

## a. Garawi's models, froma cross sectional to a panel-based analysis

## b. Decomposing the GII


# 4. Predictive modelling


# 5. Conclusions


# 6. References


# 7. Appendix
